Query Optimization for Dynamic Imputation

### 2017-03-01 paper/algorithm.tex

\subsection{Imputation operators}
\label{sec:operators}
We introduce two new relational operators to perform imputation: \textit{Impute} ($\mu$) and
\textit{Drop} ($\delta$). Each operator takes arguments $(C, R)$ where $C$ is a set of
attributes and $R$ is a relation. \textit{Impute} uses a machine learning algorithm to
replace all \nullv{} values with non-\nullv{} values for attributes $C$ in the relation $R$ (discussed in detail in~\Cref{sec:imputation}).
In our experiments, we instantiate \textit{Impute} with the CE-CART~\cite{burgette2010multiple} algorithm, but we highlight that
our system is agnostic to this choice.
\textit{Drop} simply removes from $R$ all tuples which have a \nullv{} value for some attribute in $C$.
Both operators guarantee that the resulting relation will contain no \nullv{} values for
attributes in $C$.

>>>>

\subsection{Imputation operators}
\label{sec:operators}
We introduce two new relational operators to perform imputation: \textit{Impute} ($\mu$) and
\textit{Drop} ($\delta$). Each operator takes arguments $(C, R)$ where $C$ is a set of
attributes and $R$ is a relation. \textit{Impute} uses a machine learning algorithm to
replace all \nullv{} values with non-\nullv{} values for attributes $C$ in the relation $R$ (discussed in detail in~\Cref{sec:imputation}).
In our experiments, we implement \textit{Impute} using a chained-equation decision trees algorithm, but our system is agnostic to this choice.
\textit{Drop} simply removes from $R$ all tuples which have a \nullv{} value for some attribute in $C$.
Both operators guarantee that the resulting relation will contain no \nullv{} values for
attributes in $C$.

### 2017-03-01 paper/main.tex
paper: cleanup and remove todo in conclusion
paper: balance last page
paper: minor formatting & reference fix

@@

\begin{abstract}
Missing values abound in data analysis and often present a serious challenge for usability of data. Users
are forced to pick between removing tuples with missing values or creating a clean duplicate of their data
by applying a potentially-expensive imputation strategy. Our system, \ProjectName{}, incorporates imputation into a cost-based query optimizer, 
performing any operations needed on the fly
for each query. This allows users to immediately explore
their data, while the system picks the optimal placement of imputation operations. This placement is based
on the specific query issued and a user-provided parameter that trades off data
quality and runtime performance. We incorporate the imputation operators into the relational algebra, allowing
for traditional optimizations to take place. Our experiments evaluate this approach on two real-world survey-based datasets
from the \textit{Center for Disease Control and Prevention} and \textit{freeCodeCamp}. Our
results show that the optimizer generates query plans that execute between \lowxalphazero{}
and \highxalphazeroexacs{} times as fast as imputing at the base tables and then executing
queries, making exploratory data analysis using our system feasible.  Furthermore, we show
that the query results from on-the-fly imputation have
Symmetric-Mean-Absolute-Percentage-Error between \lowsmapealphazero{} and
\highsmapealphaoneexacs{} percent when compared to standard full-table imputation
approach.

@@

\input{introduction}
\input{examples}
\input{algorithm}
\input{experiments}
\input{relatedwork}
\input{conclusion}
\bibstyle{abbrv}
\printbibliography
\balance

>>>>

\hyphenation{Im-pute-DB}

@@

\begin{abstract}
Missing values abound in data analysis and often present a serious challenge for the usability of data. Users
are forced to pick between removing tuples with missing values or creating a clean duplicate of their data
by applying a potentially-expensive imputation strategy. Our system, \ProjectName{}, incorporates imputation into a cost-based query optimizer, 
performing any operations needed on the fly
for each query. This allows users to immediately explore
their data, while the system picks the optimal placement of imputation operations. This placement is based
on the specific query issued and a user-provided parameter that trades off data
quality and runtime performance. We incorporate the imputation operators into the relational algebra, allowing
for traditional optimizations to take place. Our experiments evaluate this approach on two real-world survey-based datasets
from the \textit{Center for Disease Control and Prevention} and \textit{freeCodeCamp}. Our
results show that the optimizer generates query plans that execute between \lowxalphazero{}
and \highxalphazeroexacs{} times as fast as imputing at the base tables and then executing
queries, making exploratory data analysis using our system feasible.  Furthermore, we show
that the query results from on-the-fly imputation have
error between \lowsmapealphazero{} and
\highsmapealphaoneexacs{} percent when compared to standard full-table imputation
approach.

@@

\input{introduction}
\input{examples}
\input{algorithm}
\input{experiments}
\input{relatedwork}
\input{conclusion}
\bibstyle{abbrv}
\balance
\printbibliography
